---
title: "Data Import"
output: github_document
---

```{r setup}
library(tidyverse)
# Excel：readxl后面才用，但写在前面比较好
library(readxl)
# SAS
library(haven)
```

## Read in some data

read in the litters dataset

```{r}

litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## take a look at the data

always take a look at your data!!!

```{r}
#查看变量名称
names(litters_df)
```

printing in the console!
View(litters_df)

```{r}
litters_df
```

```{r}
head(litters_df)
tail(litters_df)
# check tail 以防数据末端异常
```

```{r}
# ::等同于library。现实中不怎么用
skimr::skim(litters_df)
# 经常用的是View(litters_df)，直接在console里写！！！
# ?read_csv可以在help里查看
```

## fix the missing data

```{r}
litters_df = 
  read.csv("./data/FAS_litters.csv" , 
           na=c("NA","O","."))
```

## import pups data

```{r}
pups_df = 
  read_csv("./data/FAS_pups.csv", 
           skip = 1,
           na=c("NA","O",".")
           )

pups_df = 
  janitor::clean_names(pups_df)

pups_df  
```

## options to read_csv

check out `?read_csv()` for more information.
```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv", 
           skip = 10, 
           col_names = FALSE)
```



## other file format

Csv file is better than excel

read in excel file

```{r}
mlb_df = 
  read_excel("./data/mlb11.xlsx",
             na=c("NA","O","."))

mlb_df

# View(mlb_df)写在console！！！

# ?read_excel写在console
```

# range

```{r}
LotR_df = 
  read_excel("./data/LotR_Words.xlsx",
             range = "B3:D6")

LotR_df
```


## SAS

read in a SAS file

```{r}
pulse_df = 
  read_sas("./data/public_pulse_data.sas7bdat")

pulse_df =
  janitor::clean_names(pulse_df)

pulse_df
```

## comparison with base R

what about `read.csv`...?
不要用read.csv！！！

```{r eval=FALSE}
litter_base = 
  read.csv("./data/FAS_litters.csv")

#用下面这个！！！
litter_readr = 
  read_csv("./data/FAS_litters.csv") 

litter_base #read.csv(_)会输出所有数据
litter_readr
```

## exporting data

export the mlb sub-table

```{r}
write_csv(LotR_df, "./data/LotR_subtable.csv")

write_csv(mlb_df, "./data/mlb_subtable.csv")
```

